---
title: 'MSCS 264: Homework #14 Key'
output: md_document
---

You should submit a knitted pdf file on Moodle, but be sure to show all of your R code, in addition to your output, plots, and written responses.

```{r, setup, include=FALSE}
library(tidyverse) 
library(ggmap)
library(stringr)
library(leaflet)
```

##Factors, Functions, and Vectors

1. Given the output below, write out the elements of the vector `y`.

```{r}
#> sort(y)
#[1] -2 -2 -1  0  3  4  4  7
#
#> order(y)
#[1] 1 8 4 5 6 2 7 3
```

```{r}
y <- c(-2, 4, 7, -1, 0, 3, 4, -2)
sort(y)
order(y)
```

2. Explain what the following function does and demonstrate by running `foo1(x)` with a few appropriately chosen vectors `x`.

```{r}
foo1 <- function(x) {
  diff <- x[-1] - x[1:(length(x) - 1)]
  length(which(diff < 0))
}
foo1(y)
```

foo1 finds the difference between successive values of `x` and counts how many of those differences are negative (e.g. drop from the 5th value to the 6th)

3. Write a function called `greet` using `if`, `else if`, and `else` to print out "good morning" if it's before 12 PM, "good afternoon" if it's between 12 PM and 5 PM, and "good evening" if it's after 5 PM.  Your function should work if you input a time like: `greet(time = "2018-05-03 17:38:01 CDT")` or if you input the current time with `greet(time = Sys.time())`.

```{r}
hours <- parse_number(str_sub(Sys.time(), 12, 13))
greet <- function(time = Sys.time()) {
  hour <- parse_number(str_sub(time, 12, 13))
  if (hour < 12) {
    print("good morning")
  } else if (hour < 17) {
    print("good afternoon")
  } else {
    print("good evening")
  }
}
greet()
greet(time = "2018-05-03 17:38:01 CDT")
```


##Airbnb Data

4. `airbnbData.csv` contains data scraped on 2022 airbnb units in the Boston area (65 variables).  The following code produces a list containing some airbnb data.  Show how to pull these data elements from `airbnb_list` (that is, your code MUST use `airbnb_list`):

- the last 10 titles of listings after sorting in alphabetical order
- the median listing price
- the latitude and longitude of listing #3774132
- the number of listings with a host named Paul
- print out the sentence "There are XX listings with wheelchair access." where XX is automatically filled in 

```{r}
airbnb.df <- read.csv("datasets/airbnbData.csv")
airbnb.df <- as.tibble(airbnb.df) 

titles <- airbnb.df$Title
eda_Price <- airbnb.df %>%
    summarise(mean = mean(Price, na.rm = TRUE),
              median = median(Price, na.rm = TRUE),
              sd = sd(Price, na.rm = TRUE),
              iqr = IQR(Price, na.rm = TRUE),
              n = n())
lat_long <- airbnb.df %>% select(ListingID, Lat, Long)
host_names <- airbnb.df %>% select(HostName) %>% distinct() %>% pull(HostName)
wheelchair <- sum(airbnb.df$A_Wheelchair)
airbnb_list <- list(titles, eda_Price, lat_long, host_names, wheelchair)

sort(airbnb_list[[1]])[2013:2022]
airbnb_list[[2]] %>% select(median)
airbnb_list[[3]] %>% filter(ListingID == 3774132)
sum(str_detect(airbnb_list[[4]], "Paul"))
paste0("There are ", airbnb_list[[5]], " listings with wheelchair access.")
```

5. The code below shows a function which returns a set of summary statistics for a single variable.  Note a couple of strange elements (e.g. `enquo` and `!!var`) that were needed to get around the way tibbles and dplyr handle variable names (for way more detail than you probably want on this issue, see https://cran.r-project.org/web/packages/dplyr/vignettes/programming.html). 

a) Create a new function called `eda2stats` which produces the same set of summary statistics by group.  Show your function works by running `eda2stats(airbnb.df, Price, S_PropType)`.

b) Run `eda2stats` again after forming a new Property Type variable with only 4 levels: Apartment, House, Bed & Breakfast, and Other.

```{r}
# First remove 5 units with prices over 1000
airbnb.df <- airbnb.df %>%
  filter(Price < 1000)

# Find summary stats on a single variable
eda1stats <- function(df, var) {
  var <- enquo(var)
  df %>%
    summarise(mean = mean(!!var, na.rm = TRUE),
              median = median(!!var, na.rm = TRUE),
              sd = sd(!!var, na.rm = TRUE),
              iqr = IQR(!!var, na.rm = TRUE),
              n = n())
}
eda1stats(airbnb.df, Price)

# Find summary stats by a grouping variable - part (a)
eda2stats <- function(df, var, groups) {
  var <- enquo(var)
  groups <- enquo(groups)
  df %>%
    group_by(!!groups) %>%
    summarise(mean = mean(!!var, na.rm = TRUE),
              median = median(!!var, na.rm = TRUE),
              sd = sd(!!var, na.rm = TRUE),
              iqr = IQR(!!var, na.rm = TRUE),
              n = n())
}
eda2stats(airbnb.df, Price, S_PropType)

# Create fewer levels of the grouping variable - part (b)
airbnb.df <- airbnb.df %>%
  mutate(S_PropType2 = fct_lump(S_PropType, n = 3))
eda2stats(airbnb.df, Price, S_PropType2)
```

6. Use `stringr` functions to extract at least one feature from `AboutListing` and create a plot to determine if that feature is related to `Price`.  A couple of examples are illustrated below.

```{r, warning=FALSE}
# Analyze features of AboutListing description
foo1 <- airbnb.df$AboutListing[5]
str_length(foo1)
str_count(foo1, "!")
str_detect(foo1, "[Ww]alk")
str_length(airbnb.df$AboutListing)[1:50]   # missings seems to have length 7

airbnb.df <- airbnb.df %>%
  mutate(NumExclaim = str_count(AboutListing, "!"),
         PctExclaim = ifelse(str_length(AboutListing) == 7, NA, 
           str_count(AboutListing, "!") / str_length(AboutListing)),
         Walk = str_detect(AboutListing, "[Ww]alk"))
airbnb.df %>% group_by(Walk) %>% tally()
airbnb.df %>% group_by(NumExclaim) %>% tally()
with(airbnb.df, summary(PctExclaim))

ggplot(airbnb.df, aes(x = Walk, y = Price)) +
  geom_boxplot() + coord_flip()
ggplot(airbnb.df, aes(x = NumExclaim, y = Price)) +
  geom_point(size = 0.25) + 
  geom_smooth(method = "lm", colour = "red")
ggplot(airbnb.df, aes(x = PctExclaim, y = Price)) +
  geom_point(size = 0.25) + 
  geom_smooth(method = "loess", colour = "red")
```


#Extra Credit 

7. Create a plot that uses a map of Boston as its backdrop and illustrates some trends in the data (assuming the packages ggmap and leaflet are loaded on the R server and ggmap is actually working).  A few examples are shown below.

```{r, warning = FALSE, message = FALSE}
## Superimpose points on street map of Boston, then size and color 
##   points in meaningful ways
qmplot(Long, Lat, data = airbnb.df, 
       colour = I('red'), size = I(1), darken = .1)  # works

bbox <- c(left = min(airbnb.df$Long), bottom = min(airbnb.df$Lat), 
          right = max(airbnb.df$Long), top = max(airbnb.df$Lat))
bostonmap <- get_stamenmap(bbox, zoom = 12, maptype = "toner-lite")
ggmap(bostonmap)
bostonmap2 <- get_stamenmap(bbox, zoom = 12, maptype = "terrain")
ggmap(bostonmap2)

ggmap(bostonmap) +
  geom_point(data = airbnb.df, aes(x = Long, y = Lat, color = S_PropType,
                                   size = parse_number(S_Accomodates))) +
  scale_size_continuous(range = c(.25, 2.5))

# Density plot of airbnb units
ggmap(bostonmap, legend = "bottomleft", extent = "device") + 
  stat_density2d(aes(x = Long, y = Lat, fill = ..level..), alpha = .5,
    data = airbnb.df, geom = "polygon")

ggmap(bostonmap) + 
  stat_density2d(aes(x = Long, y = Lat, fill = ..level..), alpha = .5,
    data = airbnb.df, geom = "polygon") +
  scale_fill_gradient(low = "black", high= "red") +
  facet_wrap(~ S_PropType2)
```

```{r, eval = FALSE}
# Interactive map with leaflet
#   addTiles()   Add background map
#   setView()    Set where the map should originally zoom to
leaflet() %>%
    addTiles() %>% 
    setView(lng = mean(airbnb.df$Long), lat = mean(airbnb.df$Lat), 
            zoom = 13) %>% 
    addCircleMarkers(data = airbnb.df,
        lat = ~ Lat, lng = ~ Long,
        radius = ~ parse_number(S_Accomodates), popup = ~ AboutListing, weight = 3,
        color = "red", fillColor = "yellow")
```
